Predicting Bike Usage at UofT St. George with Bayesian Poisson Regression

### 2024-11-29 paper/paper.qmd
measurement updated
measurement updated

Many bike-sharing service providers worldwide have made their operational data publicly available for research purposes. For instance, the Shenzhen Municipal Transport Bureau in China released a dataset containing partial bike-sharing order information from January to August 2021. This dataset includes eight variables: user ID, start time, start longitude, start latitude, end time, end longitude, end latitude, and company ID, with a total of approximately 240 million records [@szdd]. While this dataset offers a rich variety of variables, it has limitations that make it unsuitable for my research. The data only covers an eight-month period, which is insufficient to capture long-term trends in bike-sharing usage. Additionally, university campuses in China are typically closed environments, unlike the open urban settings of Toronto, making the data less generalizable for my research context. Similar issues are present in other bike-sharing datasets from different regions. These include limited temporal coverage, which restricts long-term analysis, strong regional characteristics influenced by local cultural habits, and significant variations in data quality and structure due to fierce market competition among bike-sharing companies. These factors make it challenging to integrate data across companies or regions. Therefore, while these datasets may hold value for other studies, they do not meet the requirements of my research.

@@

To predict the relationship between bike-sharing usage and time within the University of Toronto's St. George campus, this study requires a time series to describe the changes in usage at different stations over time. Therefore, the dataset published by opendatatoronto [@opendata], which records every instance of bike-sharing usage in the Toronto area since 2015, is ideal for this study. Additionally, this data is ideal because of its reliability—due to the commercial nature of bike-sharing, the specific time and location of each bike's use and return are accurately recorded. However, the raw data cannot be used directly in this project; sophisticated data cleaning is required, with specific steps and reasons as follows:

First, data from 2015-2016 was excluded because the collection and recording methods for those years differ significantly from later years and do not include the time and station variables required for this study. Next, data for stations located within the University of Toronto's St. George campus was extracted from all remaining samples, and the usage of each station was calculated for every four-hour interval. Finally, the variables of interest were extracted, and the data format was standardized for subsequent analysis. The cleaned data is presented in the format of @tbl-cleaned_data.

>>>>

There are many similar datasets available, but they pose significant limitations in the context of this study. Bike-sharing service providers worldwide have made their operational data publicly available for research purposes. For instance, the Shenzhen Municipal Transport Bureau in China released a dataset containing partial bike-sharing order information from January to August 2021. This dataset includes eight variables: user ID, start time, start longitude, start latitude, end time, end longitude, end latitude, and company ID, with a total of approximately 240 million records [@szdd]. While this dataset offers a rich variety of variables, it has limitations that make it unsuitable for this research. The data only covers an eight-month period, which is insufficient to capture long-term trends in bike-sharing usage. Additionally, university campuses in China are typically closed environments, unlike the open urban settings of Toronto, making the data less generalizable for this study's research context. Similar issues are present in other bike-sharing datasets from different regions. These include limited temporal coverage, which restricts long-term analysis, strong regional characteristics influenced by local cultural habits, and significant variations in data quality and structure due to fierce market competition among bike-sharing companies. These factors make it challenging to integrate data across companies or regions. Therefore, while these datasets may hold value for other studies, they do not meet the requirements of this research.

@@

To predict the relationship between bike-sharing usage and time within the University of Toronto's St. George campus, this study requires a time series to describe the changes in usage at different stations over time. Therefore, the "Bike Share Toronto Ridership Data" [@bikesharedata], which spans over eight years to date and documents every instance of bike-sharing usage, including the time and location of each ride, is ideal for this study.

This data is particularly ideal for the study also due to its high reliability, which stems from the commercial nature of bike-sharing operations. The advanced bike-sharing system ensures accurate data collection and management by automatically recording ride information. Each bike and docking station is equipped with sensors and communication devices, enabling the capture of detailed trip data.

However, the raw data cannot be used directly in this project; sophisticated data cleaning is required, with specific steps and reasons as follows: First, data from 2015-2016 was excluded because the collection and recording methods for those years differ significantly from later years and do not include the time and station variables required for this study. Next, data for stations located within the University of Toronto's St. George campus was extracted from all remaining samples, and the usage of each station was calculated for every four-hour interval. Finally, the variables of interest were extracted, and the data format was standardized for subsequent analysis. The cleaned data is presented in the format of @tbl-cleaned_data.
